---
title: "Homework 3"
author: "Ruoyuan Qian"
output:   
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
#install.packages("Rmisc")
#install.packages("plyr")
#install.packages("DescTools")
library(tidyverse)
library(ggplot2)
library(plyr)
library(readxl)
library(dplyr)
library(arsenal)
library(ggpubr)
library(Rmisc)
library(DescTools)

```

# Problem 2

## a)

Input data
```{r}
surv = read_csv(".\\data\\SurvCancer.csv")
```

Make discriptive statistics table
```{r}
surv %>% 
  group_by(ORGAN) %>% 
  mutate( N = n(),
          Mean = round(mean(SURVIVAL),2),
          SD = round(sd(SURVIVAL),2),
          Median = median(SURVIVAL),
          IQR = str_c(quantile(SURVIVAL)[2],"-",
                      quantile(SURVIVAL)[4]),
          Range = str_c(min(SURVIVAL),"-",max(SURVIVAL))) %>% 
  select(-SURVIVAL) %>% 
  distinct() %>% 
  knitr::kable(caption = "Table 1  Discriptive statistics of each group",
               format = "html") 


```


Make boxplot of survival time
```{r}
boxplot(SURVIVAL~ORGAN,data=surv,main="Survival time in each organ group",xlab="Organ", ylab="Survival time")

```


Make density plot 
```{r}
Stomach_his = 
surv %>% 
filter(ORGAN=="Stomach") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 140)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Stomach")

Bronchus_his = 
surv %>% 
filter(ORGAN=="Bronchus") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 130)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Bronchus")

Colon_his = 
surv %>% 
filter(ORGAN=="Colon") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 140)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Colon")

Ovary_his = 
surv %>% 
filter(ORGAN=="Ovary") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 400)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Ovary")

Breast_his = 
surv %>% 
filter(ORGAN=="Breast") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 400)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Breast")

multiplot(Stomach_his,Bronchus_his,
          Colon_his,Ovary_his,Breast_his,
          cols=2)
```


## b)

ANOVA 

Calculated by R
```{r}
res1<-aov(SURVIVAL~factor(ORGAN), data=surv)

summary(res1) 
```

Calculated by hand
```{r}

n = surv %>% 
  pull(ORGAN) %>% 
  length()

k = surv %>% 
  distinct(ORGAN) %>% 
  pull(ORGAN) %>% 
  length()

total_ss = sum((surv[,1])^2)-(sum(surv[,1]))^2/n

sum_y2 = (sum(surv[,1]))^2/n

between_ss = 
surv %>% 
  group_by(ORGAN) %>% 
  mutate(n = n(),
            mean = mean(SURVIVAL)) %>% 
  distinct(mean,n) %>% 
  ungroup(ORGAN) %>% 
  mutate(mean^2*n,
         sum(mean^2*n),
         between_ss = sum(mean^2*n)-sum_y2) %>% 
  select(between_ss) %>% 
  distinct() %>% 
  pull(between_ss) 

whithin_ss = total_ss - between_ss

MSB = between_ss/(k-1)
MSE = whithin_ss/(n-k)

f_s = MSB/MSE

p_value = 1-pf(f_s,(k-1),(n-k))

f_a = qf(0.99,4,59)

f_s > f_a
```

## c)

Pairwise comparison
```{r}
pairwise.t.test(surv$SURVIVAL, surv$ORGAN, p.adj='bonferroni')

TukeyHSD(res1,conf.level = 0.99)

DunnettTest(surv$SURVIVAL, factor(surv$ORGAN))
```


## d)

Make denrity and Q-Q plot
```{r}
Stomach =
 surv %>% 
  filter(ORGAN == "Stomach") %>% 
  pull(SURVIVAL)

Bronchus =
 surv %>% 
  filter(ORGAN == "Bronchus") %>% 
  pull(SURVIVAL)

Colon =
 surv %>% 
  filter(ORGAN == "Colon") %>% 
  pull(SURVIVAL)

Ovary =
 surv %>% 
  filter(ORGAN == "Ovary") %>% 
  pull(SURVIVAL)

Breast =
 surv %>% 
  filter(ORGAN == "Breast") %>% 
  pull(SURVIVAL)

```
```{r}
Stomach_his = 
surv %>% 
filter(ORGAN=="Stomach") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 140)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Stomach")

Stomach_qq = 
ggqqplot(Stomach,main="Q-Q plot for intervention group")


Bronchus_his = 
surv %>% 
filter(ORGAN=="Bronchus") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 130)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Bronchus")

Bronchus_qq = 
ggqqplot(Bronchus,main="Q-Q plot for intervention group")


Colon_his = 
surv %>% 
filter(ORGAN=="Colon") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 140)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Colon")

Colon_qq = 
ggqqplot(Colon,main="Q-Q plot for intervention group")


Ovary_his = 
surv %>% 
filter(ORGAN=="Ovary") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 400)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Ovary")

Ovary_qq = 
ggqqplot(Ovary,main="Q-Q plot for intervention group")


Breast_his = 
surv %>% 
filter(ORGAN=="Breast") %>% 
 ggplot(aes(SURVIVAL,..density..))+
 geom_histogram(binwidth = 400)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Breast")

Breast_qq = 
ggqqplot(Breast,main="Q-Q plot for intervention group")



multiplot(Stomach_his,Bronchus_his,
          Stomach_qq,Bronchus_qq, 
          cols=2)

multiplot(
          Colon_his,Ovary_his,Breast_his,
          Colon_qq,Ovary_qq, Breast_qq,
          cols=2)

```

Perform Shapiro-Wilk test
```{r}
shapiro.test(Stomach)
shapiro.test(Bronchus)
shapiro.test(Colon)
shapiro.test(Ovary)
shapiro.test(Breast)
```
### i)

Transformation can be done to tranaform the data into normality.

Krusak-Wallis test can be used when the normality is questionable.

### ii)

Log-transformation
```{r}
surv_log = 
  surv %>% 
  mutate(log = log(SURVIVAL))

```

Make density and Q-Q plots after log-transformation
```{r}
Stomach =
 surv_log %>% 
  filter(ORGAN == "Stomach") %>% 
  pull(log)

Bronchus =
 surv_log %>% 
  filter(ORGAN == "Bronchus") %>% 
  pull(log)

Colon =
 surv_log %>% 
  filter(ORGAN == "Colon") %>% 
  pull(log)

Ovary =
 surv_log %>% 
  filter(ORGAN == "Ovary") %>% 
  pull(log)

Breast =
 surv_log %>% 
  filter(ORGAN == "Breast") %>% 
  pull(log)

```

```{r}
Stomach_his = 
surv_log %>% 
filter(ORGAN=="Stomach") %>% 
 ggplot(aes(log,..density..))+
 geom_histogram(binwidth = .5)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Stomach")

Stomach_qq = 
ggqqplot(Stomach,main="Q-Q plot for intervention group")


Bronchus_his = 
surv_log %>% 
filter(ORGAN=="Bronchus") %>% 
 ggplot(aes(log,..density..))+
 geom_histogram()+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Bronchus")

Bronchus_qq = 
ggqqplot(Bronchus,main="Q-Q plot for intervention group")


Colon_his = 
surv_log %>% 
filter(ORGAN=="Colon") %>% 
 ggplot(aes(log,..density..))+
 geom_histogram()+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Colon")

Colon_qq = 
ggqqplot(Colon,main="Q-Q plot for intervention group")


Ovary_his = 
surv_log %>% 
filter(ORGAN=="Ovary") %>% 
 ggplot(aes(log,..density..))+
 geom_histogram(binwidth = 0.6)+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Ovary")

Ovary_qq = 
ggqqplot(Ovary,main="Q-Q plot for intervention group")


Breast_his = 
surv_log %>% 
filter(ORGAN=="Breast") %>% 
 ggplot(aes(log,..density..))+
 geom_histogram()+
 geom_line(stat = 'density',size = 1)+
    labs(x = "Breast")

Breast_qq = 
ggqqplot(Breast,main="Q-Q plot for intervention group")



multiplot(Stomach_his,Bronchus_his,
          Stomach_qq,Bronchus_qq, 
          cols=2)

multiplot(
          Colon_his,Ovary_his,Breast_his,
          Colon_qq,Ovary_qq, Breast_qq,
          cols=2)

```

Perform Shapiro-Wilk test
```{r}
shapiro.test(Stomach)
shapiro.test(Bronchus)
shapiro.test(Colon)
shapiro.test(Ovary)
shapiro.test(Breast)
```

KW.test for non-normality data
```{r}
kruskal.test(surv$SURVIVAL~surv$ORGAN)

res1<-aov(SURVIVAL~factor(ORGAN), data=surv)

summary(res1) 
```

# Problem 3

## a)

group nz: placebo and iron 

group z: zinc and zinc+ iron

group pla: placebo

group iro: iron

group z: zinc

group zi: zinc+iron

Input data
```{r}
treatment = c("placebo","iron","zinc","zinc+iron")
n = c(56,54,54,55)
mean = c(1.1,1.4,0.7,0.8)
se = c(0.2,0.2,0.1,0.1)

data = tibble(treatment,n,mean,se)

data_nz = 
  data %>% 
  filter(treatment %in% c("placebo","iron"))

data_z = 
  data %>% 
  filter(treatment %in% c("zinc","zinc+iron"))

```

Calculate the averages and difference in the averages
```{r}
mean_nz =
  data_nz %>% 
  mutate(new_mean = sum(n * mean)/sum(n)) %>% 
  select(new_mean) %>% 
  distinct %>% 
  pull(new_mean)

mean_z =
  data_z %>% 
  mutate(new_mean = sum(n * mean)/sum(n)) %>% 
  select(new_mean) %>% 
  distinct %>% 
  pull(new_mean)

n_z = 
  data_z %>% 
  mutate(n_z = sum(n)) %>% 
  select(n_z) %>% 
  distinct %>% 
  pull(n_z)

n_nz = 
  data_nz %>% 
  mutate(n_nz = sum(n)) %>% 
  select(n_nz) %>% 
  distinct %>% 
  pull(n_nz)

se_nz = 
  data_nz %>% 
  mutate(new_se = sqrt(sum(se^2))) %>% 
  select(new_se) %>% 
  distinct %>% 
  pull(new_se)

se_z = 
  data_z %>% 
  mutate(new_se = sqrt(sum(se^2))) %>% 
  select(new_se) %>% 
  distinct %>% 
  pull(new_se)

d1 = round(mean_nz - mean_z,2)
  
```

## b)

Calculate sd for each original group
```{r}
data_sd = 
data %>% 
  mutate(sd = sqrt(se^2*n)) %>% 
  as.matrix()
```

F-test for ‘zinc’ and ‘zinc-iron' group
```{r }

sd_pla = as.numeric(data_sd[1,5])
sd_iro = as.numeric(data_sd[2,5])

n_pla = as.numeric(data_sd[1,2])
n_iro = as.numeric(data_sd[2,2])

            
f_s = sd_pla^2/sd_iro^2

f_a = qf(0.975,n_pla-1,n_iro-1)

f_s < f_a
```

So sd is equal between placebo and iron.

F-test for ‘placebo’ and ‘iron’ group
```{r}
sd_z = as.numeric(data_sd[3,5])
sd_zi = as.numeric(data_sd[4,5])

n_z = as.numeric(data_sd[3,2])
n_zi = as.numeric(data_sd[4,2])

f_s = sd_zi^2/sd_z^2

f_a = qf(0.975,n_zi-1,n_z-1)

f_s < f_a
```

So sd is equal between zinc and zinc+iron.

Calculate the var for new groups
```{r}
s_nz_2 = round(((n_pla-1)*sd_pla^2+(n_iro-1)*sd_iro^2)/(n_pla+n_iro-2),2)

s_z_2 = round(((n_z-1)*sd_z^2+(n_zi-1)*sd_zi^2)/(n_z+n_zi-2),2)
```

## c)

### i)

Sample size for equal allocation
```{r}
k = 1
n1 = (s_nz_2+s_z_2)*(qnorm(0.975)+qnorm(0.8))^2/d1^2
n2 = (s_nz_2+s_z_2)*(qnorm(0.975)+qnorm(0.8))^2/d1^2

```

Sample size for 2:1 allocation
```{r}
k = 2
n1 = (s_nz_2+s_z_2/0.5)*(qnorm(0.975)+qnorm(0.8))^2/d1^2
n2 = (s_nz_2*0.5+s_z_2)*(qnorm(0.975)+qnorm(0.8))^2/d1^2
```


